---
title: "Intensity_IHS_analysis"
author: "Franziska Hildebrandt"
date: "2023-11-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First load the intensity data for 1. raw data and 2. binarized data

```{r}

raw <- read.table("/Users/franziskahildebrandt/Downloads/Raw_Data_Hotspots_IHSraw.csv", header = T, dec = ".", check.names = T, sep = ";")
raw <- raw[,1:13]
head(raw)

binary <- read.table("/Users/franziskahildebrandt/Downloads/raw_Data_Hotspots_IHSbinary_csv.csv", header = T, dec = ".", check.names = T, sep = ";")
binary <- binary[,1:13]
head(binary)

library(dplyr)

##remove IBA1 as the staining was incorrect + remove all NA

raw <- raw[raw$Antigen != "IBA1",]
binary <- binary[binary$Antigen != "IBA1",]

raw <- na.omit(raw)
raw$Mean <- as.numeric(raw$Mean)
binary <- na.omit(binary)

head(raw)
head(binary)

#Change mouse C to Control 24h

raw[raw$Sample %in% "200225_C",]$Timepoint <- gsub(x=raw[raw$Sample %in% "200225_C",]$Timepoint, pattern = "24h", replacement = "24h_C")
raw

binary[binary$Sample %in% "200225_C",]$Timepoint <- gsub(x=binary[binary$Sample %in% "200225_C",]$Timepoint, pattern = "24h", replacement = "24h_C")
binary

#Change mouse I to Control 38h
raw[raw$Sample %in% "200303_I",]$Timepoint <- gsub(x=raw[raw$Sample %in% "200303_I",]$Timepoint, pattern = "38h", replacement = "38h_C")


binary[binary$Sample %in% "200303_I",]$Timepoint <- gsub(x=binary[binary$Sample %in% "200303_I",]$Timepoint, pattern = "38h", replacement = "38h_C")


```



Next we scale the mean intensities (AU) to fit value between 0 and 1 

```{r,}

raw <- raw %>%
  mutate(scaled = (Mean - min(Mean)) / (max(Mean) - min(Mean)))

binary <- binary %>%
  mutate(scaled = (Mean - min(Mean)) / (max(Mean) - min(Mean)))

raw 
binary

```


Now we can split the data by timepoint
```{r}

raw_12 <- raw[raw$Timepoint %in% "12h",]
raw_24 <- raw[raw$Timepoint %in% "24h",]
raw_38 <- raw[raw$Timepoint %in% "38h",]
raw_24C <- raw[raw$Timepoint %in% "24_C",]
raw_38C <- raw[raw$Timepoint %in% "38_C",]

binary_12 <- binary[binary$Timepoint %in% "12h",]
binary_24 <- binary[binary$Timepoint %in% "24h",]
binary_38 <- binary[binary$Timepoint %in% "38h",]
binary_24C <- binary[binary$Timepoint %in% "24h_C",]
binary_38C <- binary[binary$Timepoint %in% "38h_C",]


```



Now we would like to plot the mean intensity of each antigen across the different hotspots comparing the area_type roi and inverse 

```{r}

p_12h_binary <- ggplot(binary_12, aes(x = Antigen, y = scaled, fill = Area_type)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, size = 0.8)+ 
  labs(title = "Mean Intensity in IHS across cell types 12hpi",
       x = "celltype marker",
       y = "mean Intensity [AU]") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), labels = c("encompassing tissue", "IHS")) +
  theme_classic() +
  coord_flip()

p_24h_binary <- ggplot(binary_24, aes(x = Antigen, y = scaled, fill = Area_type)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, size = 0.8)+ 
  labs(title = "Mean Intensity in IHS across cell types 24hpi",
       x = "celltype marker",
       y = "mean Intensity [AU]") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), labels = c("encompassing tissue", "IHS")) +
  theme_classic() +
  coord_flip()

p_38h_binary <- ggplot(binary_38, aes(x = Antigen, y = Mean, fill = Area_type)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, size = 0.8)+ 
  labs(title = "Mean Intensity in IHS across cell types 38hpi",
       x = "celltype marker",
       y = "mean Intensity [AU]") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), labels = c("encompassing tissue", "IHS")) +
  theme_classic() +
  coord_flip()



p_12h_binary

p_24h_binary

p_38h_binary

p_24C_binary <- ggplot(binary_24C, aes(x = Antigen, y = scaled, fill = Area_type)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, size = 0.8)+ 
  labs(title = "Mean Intensity in IHS across cell types 24h control",
       x = "celltype marker",
       y = "mean Intensity [AU]") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), labels = c("encompassing tissue", "IHS")) +
  theme_classic() +
  coord_flip()

p_38C_binary <- ggplot(binary_38C, aes(x = Antigen, y = scaled, fill = Area_type)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, size = 0.8)+ 
  labs(title = "Mean Intensity in IHS across cell types 38h control",
       x = "celltype marker",
       y = "mean Intensity [AU]") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), labels = c("encompassing tissue", "IHS")) +
  theme_classic() +
 
  
  coord_flip()

p_24C_binary

p_38C_binary

```

Now we would like to calculate wether these observed differences are significant and display significance in the plot. 


```{r, plot for 12hpi}

library(ggpubr)
library(rstatix)

stat.test <- binary_12 %>%
  group_by(Antigen) %>%
  wilcox_test(scaled ~ Area_type) %>%
  adjust_pvalue() %>%
  add_significance("p.adj")
stat.test

library(ggpubr)

# Your existing code for creating the boxplot
bxp <- ggboxplot(binary_12, x = "Antigen", y = "scaled", fill = "Area_type", palette = ) +
  labs(x = "cell type marker", y = "binarized intensity [AU]") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), name = "Area type", labels = c("surrounding tissue", "IHS")) +
  geom_point(aes(fill = Area_type), position = position_jitter(width = 0.2), alpha = 0.5)

# Create a new data frame for p-value annotations
stat.test <- stat.test %>% add_xy_position(x = "Antigen")
pvalue_labels <- data.frame(x = stat.test$x, y = rep(1.3, nrow(stat.test)), label = stat.test$p.adj.signif)

# Add p-value annotations using geom_text
pdf("/Users/franziskahildebrandt/P.Berghei-Liver/res/figures/supplementary/081123_binarized_boxplot_12h.pdf", width = 8, height = 4)
bxp+ 
  geom_text(data = pvalue_labels, aes(x = x, y = y, label = label), vjust = 0, hjust = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))
dev.off()
```


```{r, plot for 24hpi}

library(ggpubr)
library(rstatix)

stat.test <- binary_24 %>%
  group_by(Antigen) %>%
  wilcox_test(scaled ~ Area_type) %>%
  adjust_pvalue() %>%
  add_significance("p.adj")
stat.test

library(ggpubr)

# Your existing code for creating the boxplot
bxp <- ggboxplot(binary_24, x = "Antigen", y = "scaled", fill = "Area_type", palette = ) +
  labs(x = "cell type marker", y = "binarized intensity [AU]") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), name = "Area type", labels = c("surrounding tissue", "IHS")) +
  geom_point(aes(fill = Area_type), position = position_jitter(width = 0.2), alpha = 0.5)

# Create a new data frame for p-value annotations
stat.test <- stat.test %>% add_xy_position(x = "Antigen")
pvalue_labels <- data.frame(x = stat.test$x, y = rep(1.3, nrow(stat.test)), label = stat.test$p.adj.signif)

# Add p-value annotations using geom_text
pdf("/Users/franziskahildebrandt/P.Berghei-Liver/res/figures/supplementary/081123_binarized_boxplot_24h.pdf", width = 8, height = 4)
bxp+ 
  geom_text(data = pvalue_labels, aes(x = x, y = y, label = label), vjust = 0, hjust = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))
dev.off()
```

```{r, plot for 38hpi}

library(ggpubr)
library(rstatix)

stat.test <- binary_38 %>%
  group_by(Antigen) %>%
  wilcox_test(scaled ~ Area_type) %>%
  adjust_pvalue() %>%
  add_significance("p.adj")
stat.test

library(ggpubr)

# Your existing code for creating the boxplot
bxp <- ggboxplot(binary_38, x = "Antigen", y = "scaled", fill = "Area_type", palette = ) +
  labs(x = "cell type marker", y = "binarized intensity [AU]") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), name = "Area type", labels = c("surrounding tissue", "IHS")) +
  geom_point(aes(fill = Area_type), position = position_jitter(width = 0.2), alpha = 0.5)

# Create a new data frame for p-value annotations
stat.test <- stat.test %>% add_xy_position(x = "Antigen")
pvalue_labels <- data.frame(x = stat.test$x, y = rep(1.3, nrow(stat.test)), label = stat.test$p.adj.signif)

# Add p-value annotations using geom_text
pdf("/Users/franziskahildebrandt/P.Berghei-Liver/res/figures/supplementary/081123_binarized_boxplot_38h.pdf", width = 8, height = 4)
bxp+ 
  geom_text(data = pvalue_labels, aes(x = x, y = y, label = label), vjust = 0, hjust = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))
dev.off()
```


```{r, plot for 24 control}

library(ggpubr)
library(rstatix)

stat.test <- binary_24C %>%
  group_by(Antigen) %>%
  wilcox_test(scaled ~ Area_type) %>%
  adjust_pvalue() %>%
  add_significance("p.adj")
stat.test

library(ggpubr)

# Your existing code for creating the boxplot
bxp <- ggboxplot(binary_24C, x = "Antigen", y = "scaled", fill = "Area_type", palette = ) +
  labs(x = "cell type marker", y = "binarized intensity [AU]") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), name = "Area type", labels = c("surrounding tissue", "IHS")) +
  geom_point(aes(fill = Area_type), position = position_jitter(width = 0.2), alpha = 0.5)

# Create a new data frame for p-value annotations
stat.test <- stat.test %>% add_xy_position(x = "Antigen")
pvalue_labels <- data.frame(x = stat.test$x, y = rep(1.3, nrow(stat.test)), label = stat.test$p.adj.signif)

# Add p-value annotations using geom_text
pdf("/Users/franziskahildebrandt/P.Berghei-Liver/res/figures/supplementary/081123_binarized_boxplot_24h_control.pdf", width = 8, height = 4)
bxp+ 
  geom_text(data = pvalue_labels, aes(x = x, y = y, label = label), vjust = 0, hjust = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))
dev.off()
```


```{r, plot for 38 control}

library(ggpubr)
library(rstatix)

binary_38C <- binary_38C[order(match(binary_38C$Antigen,c("CD8", "CD4", "DAPI", "F4/80", "CD27", "CD11c", "CD11b") )),]

stat.test <- binary_38C %>%
  group_by(Antigen) %>%
  wilcox_test(scaled ~ Area_type) %>%
  adjust_pvalue() %>%
  add_significance("p.adj")
stat.test

library(ggpubr)

# Your existing code for creating the boxplot
bxp <- ggboxplot(binary_38C,x = "Antigen", 
               y = "scaled", fill = "Area_type") + 
  labs(x = "cell type marker", y = "binarized intensity [AU]") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("cadetblue", "darkorange"), name = "Area type", labels = c("surrounding tissue", "IHS")) +
  geom_point(aes(fill = Area_type), position = position_jitter(width = 0.2), alpha = 0.5)

# Create a new data frame for p-value annotations
stat.test <- stat.test %>% add_xy_position(x = "Antigen")
pvalue_labels <- data.frame(x = stat.test$x, y = rep(1.3, nrow(stat.test)), label = stat.test$p.adj.signif)

# Add p-value annotations using geom_text
pdf("/Users/franziskahildebrandt/P.Berghei-Liver/res/figures/supplementary/081123_binarized_boxplot_38h_control.pdf", width =8, height = 4)
bxp+ 
  geom_text(data = pvalue_labels, aes(x = x, y = y, label = label), vjust = 0, hjust = 0.5, size = 5) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1)))
dev.off()

```